A Survey of Real-Time Strategy Game AI\\ Research and Competition in StarCraft

### 2013-03-09 survey.tex
Some revisions on Section III.A and III.B

Systems that play RTS games need to address most, if not all, the aforementioned problems together. Therefore, it is hard to classify existing work on RTS AI as addressing the different problems above. For that reason, in order to classify existing work on RTS AI, we will divide the existing work according to three levels of abstraction (division that is widely used by RTS players): strategy, tactics and reactive control (micro-management).

@@

Following this idea, we consider strategy to be everything related to the technology trees, build-order\footnote{The {\em build-order} is the specific sequence in which buildings of different types will be constructed at the beginning of a game, and completely determines the long-term strategy of a player.}, upgrades, and army composition. It is the most deliberative level, as a player selected and performs a strategy with future stances (aggressive, defensive, economy, technology) and tactics in mind. We consider tactics, everything related to confrontations between groups of units. Tactical reasoning involves both spatial (exploiting the terrain) and temporal (army movements) reasoning, constrained on the possible types of attacks by the army composition of the player and their opponent. Finally, we will call reactive control (also known as micro-management) to how the player controls individual units to maximize their efficiency (when executing tactics) in real-time and adversary conditions. The main difference between tactics and reactive control is that tactical reasoning typically involves some sort of planning ahead for some short spans of time, whereas reactive control involves no planning ahead whatsoever.

@@

% INTRO:
Strategic decision making in real-time domains is still an open problem. In the context of RTS games is has been addressed using many AI techniques, like hard-coded approaches, planning-based approaches, or machine learning-based approaches. Let us cover each of these approaches in turn.

% HARD-CODED:
Hard-coded approaches have been extensively used in commercial RTS games. The most common approaches use finite state machines (FSM) \cite{FSM_AIGameProgWisdom2003} in order to let the AI author hard-code the strategy that the AI will employ. The idea behind FSMs is to decompose the AI behavior into easily manageable states, such as ``attacking'', ``gathering resources'' or ``repairing'' and establish the conditions that trigger transitions between them. Commercial approaches also include Hierarchical FSMs, in which FSMs are composed hierarchically. These hard-coded approaches have achieved a significant amount of success, and, as we will discuss later, have also been used in many academic RTS AI research systems, as discussed in Section \ref{sec:bot}. However, strategic decision making is a hard problem, and these hard-coded approaches struggle to encode dynamic, adaptive behaviors.

@@

% PLANNING:
Approaches using planning techniques have also been explored in the literature. For example Onta\~{n}\'{o}n et al. \cite{CBR_Planning} explored the use of real-time case-based planning (CBP) in the domain of Wargus (a Warcraft II clone). In their work, they used human demonstration to learn plans, that are then composed at run-time in order to form full-fledges strategies to play the game. In \cite{PlanRetrieval} they improve over  their previous CBP approach by using situation assessment for improving the quality and speed of plan retrieval. Hierarchical Task-Network (HTN) planning has also been explored with some success in the context of simpler first-person shooter games. Planning approaches offer more adaptivity of the AI strategy compared to hard-coded approaches. However, the real-time constraints of RTS games limit the planning approaches that can be applied, being HTN and case-based planning the only ones explored so far. Moreover, none of these approaches addresses any timing or scheduling issues, which are key in RTS games. On notable exception is the work of Churchill and Buro \cite{churchill2011build}, who used planning in order to construct its economic build-orders taking into account timing constraints of the different actions.

% MACHINE LEARNING:
Concerning machine learning-based approaches, Weber and Mateas \cite{WeberCig09} proposed ``a data mining approach to strategy prediction'' and performed supervised learning (from buildings features) on labeled StarCraft replays. Dereszynski et al. \cite{HMMstrat_RTS_AIIDE11} used HMM to learn the transition probabilities of sequences of constructions and kept the most probables to produce probabilistic behavior models (in StarCraft). Synnaeve and Bessi\`{e}re \cite{SynnaeveOpeningCig11} used the dataset of \cite{WeberCig09} and presented a Bayesian semi-supervised model to learn from replays and predict openings (early game strategies) from StarCraft replays. The openings are labeled by EM clustering considering appropriate features. Then, in \cite{SynnaeveAIIDE11}, they presented an unsupervised learning Bayesian model for tech-tree prediction, still using replays. % Finally \cite{GemineImitative} used the same ``build-order features'' approach to learn to adapt the bot's build order to the opponents strategy in StarCraft 2.
Finally, evolutionary approaches to determine priorities of high level tasks was explored by Young and Hawes in their QUORUM system \cite{young2012evolutionary}, showing improvement over static priorities.

% CBR:
Also falling into the machine-learning category, a significant group of researchers has explored case-based reasoning (CBR) \cite{Aamodt94CBR} approaches for strategic decision making. For example Aha et al. \cite{LTW} used CBR to perform dynamic plan retrieval in the Wargus domain. Hsieh and Sun \cite{HsiehS08} based their work on \cite{LTW} CBR model and used StarCraft replays to construct states and building sequences (``build orders''). Finally, Schadd et al. \cite{SchaddBS07} applied a CBR approach to opponent modeling through hierarchically structured models of the opponent behavior and they applied their work to the Spring RTS (Total Annihilation clone).

% SCOUTING:
One final consideration is that RTS games are typically partially observable. For example games like StarCraft implement the ``fog of war'' idea, which basically means that a player can only see the areas of the map close to her own units. Areas of the map away from the field of view of individual units are not observable. Therefore, in order to play an RTS game, players need to scout the opponent in order to obtain information about the opponent's strategy. Very few of the previous approaches deal with this problem, and use perfect information all the time (in the case of commercial games, most AI implementations cheat, since they have perfect information). In commercial games, in order to make the human player believe the built-in AI of this games does scouting, sometimes they simulate some scouting tasks as Bob Fitch described in his AIIDE 2011 keynote for the WarCraft game series and StarCraft: Broodwar. A notable exception is the work of Weber et al. \cite{WeberAIIDE11}, who used a particle model with a linear trajectory update to track opponent units under fog of war in StarCraft. They also produced tactical goals through reactive planning and goal-driven autonomy \cite{WeberCig10,Weber10}, finding the more relevant goal(s) to spawn in unforeseen situations.

%{\color{ForestGreen}
%Nobody had described an efficient way to scout the opponent... better opponent information better opponent modeling...
%}

>>>>

Systems that play RTS games need to address most, if not all, the aforementioned problems together. Therefore, it is hard to classify existing work on RTS AI as addressing the different problems above. For that reason, in order to classify existing work on RTS AI, we will divide it according to three levels of abstraction: strategy (which loosely corresponds to ``macro''), tactics and reactive control (which loosely corresponds to ``micro'').

@@

Following this idea, we consider strategy to be everything related to the technology trees, build-order\footnote{The {\em build-order} is the specific sequence in which buildings of different types will be constructed at the beginning of a game, and completely determines the long-term strategy of a player.}, upgrades, and army composition. It is the most deliberative level, as a player selects and performs a strategy with future stances (aggressive, defensive, economy, technology) and tactics in mind. We consider tactics to be everything related to confrontations between groups of units. Tactical reasoning involves both spatial (exploiting the terrain) and temporal (army movements) reasoning, constrained on the possible types of attacks by the army composition of the player and their opponent. Finally, we will call reactive control to how the player controls individual units to maximize their efficiency (when executing tactics) in real-time. The main difference between tactics and reactive control is that tactical reasoning typically involves some sort of planning ahead for some short spans of time, whereas reactive control involves no planning ahead whatsoever.

@@

% INTRO:
Strategic decision making in real-time domains is still an open problem. In the context of RTS games is has been addressed using many AI techniques, like hard-coded approaches, planning-based approaches, or machine learning-based approaches. We cover each of these approaches in turn.

% HARD-CODED:
Hard-coded approaches have been extensively used in commercial RTS games. The most common approaches use finite state machines (FSM) \cite{FSM_AIGameProgWisdom2003} in order to let the AI author hard-code the strategy that the AI will employ. The idea behind FSMs is to decompose the AI behavior into easily manageable states, such as ``attacking'', ``gathering resources'' or ``repairing'' and establish the conditions that trigger transitions between them. Commercial approaches also include Hierarchical FSMs, in which FSMs are composed hierarchically. These hard-coded approaches have achieved a significant amount of success, and, as we will discuss later, have also been used in many academic RTS AI research systems, as discussed in Section \ref{sec:bot}. However, these hard-coded approaches struggle to encode dynamic, adaptive behaviors, and are easily exploitable by adaptive opponents.

@@

% PLANNING:
Approaches using planning techniques have also been explored in the literature. For example Onta\~{n}\'{o}n et al. \cite{CBR_Planning} explored the use of real-time case-based planning (CBP) in the domain of Wargus (a Warcraft II clone). In their work, they used human demonstration to learn plans, that are then composed at run-time in order to form full-fledges strategies to play the game. In \cite{PlanRetrieval} they improve over  their previous CBP approach by using situation assessment for improving the quality and speed of plan retrieval. Hierarchical Task-Network (HTN) planning has also been explored with some success in the context of simpler first-person shooter games \cite{HTNPlanning}. Planning approaches offer more adaptivity of the AI strategy compared to hard-coded approaches. However, the real-time constraints of RTS games limit the planning approaches that can be applied, being HTN and case-based planning the only ones explored so far. Moreover, none of these approaches addresses any timing or scheduling issues, which are key in RTS games. On notable exception is the work of Churchill and Buro \cite{churchill2011build}, who used planning in order to construct its economic build-orders taking into account timing constraints of the different actions.

% MACHINE LEARNING:
Concerning machine learning-based approaches, Weber and Mateas \cite{WeberCig09} proposed a data mining approach to strategy prediction and performed supervised learning on labeled StarCraft replays. Dereszynski et al. \cite{HMMstrat_RTS_AIIDE11} used Hidden Markov Models (HMM) to learn the transition probabilities of sequences of building construction orders and kept the most probable ones to produce probabilistic behavior models (in StarCraft). Synnaeve and Bessi\`{e}re \cite{SynnaeveOpeningCig11} used the dataset of \cite{WeberCig09} and presented a Bayesian semi-supervised model to learn from replays and predict openings (early game strategies) from StarCraft replays. The openings are labeled by EM clustering considering appropriate features. Then, in \cite{SynnaeveAIIDE11}, they presented an unsupervised learning Bayesian model for tech-tree prediction, still using replays. % Finally \cite{GemineImitative} used the same ``build-order features'' approach to learn to adapt the bot's build order to the opponents strategy in StarCraft 2.
Finally, evolutionary approaches to determine priorities of high level tasks was explored by Young and Hawes in their QUORUM system \cite{young2012evolutionary}, showing improvement over static priorities.

% CBR:
Also falling into the machine-learning category, a significant group of researchers has explored case-based reasoning (CBR) \cite{Aamodt94CBR} approaches for strategic decision making. For example Aha et al. \cite{LTW} used CBR to perform dynamic plan retrieval in the Wargus domain. Hsieh and Sun \cite{HsiehS08} based their work on Aha et al.'s CBR model \cite{LTW} and used StarCraft replays to construct states and building sequences (``build orders''). Finally, Schadd et al. \cite{SchaddBS07} applied a CBR approach to opponent modeling through hierarchically structured models of the opponent behavior and they applied their work to the Spring RTS game (a ``Total Annihilation'' clone).

% SCOUTING:
One final consideration concerning strategy is that RTS games are typically partially observable. Games like StarCraft implement the ``fog of war'' idea, which basically means that a player can only see the areas of the map close to her own units. Areas of the map away from the field of view of individual units are not observable. Players need to scout in order to obtain information about the opponent's strategy. The size of the state space in StarCraft prevents solutions based on POMDPs from being directly applicable, and very few of the previous approaches deal with this problem. Most research in RTS game AI assumes perfect information all the time. In the case of commercial games, most AI implementations cheat, since they have perfect information, while the human player does not. In order to make the human player believe the AI of this games does not cheat, sometimes they simulate some scouting tasks as Bob Fitch described in his AIIDE 2011 keynote for the WarCraft and StarCraft game series. A notable exception is the work of Weber et al. \cite{WeberAIIDE11}, who used a particle model with a linear trajectory update to track opponent units under fog of war in StarCraft. They also produced tactical goals through reactive planning and goal-driven autonomy \cite{WeberCig10,Weber10}, finding the more relevant goal(s) to spawn in unforeseen situations.
